---
title: "Tripod_05g_EstabYrFreq"
author: "CaitLittlef"
date: "January 29, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---



If I'm seeking mode of establishment years for each site, I can just take mode of trees sampled at each site. I don't need to use any expansion factor because I'm not expressing on a per area basis.

From Urza & Sibold 2013, we know that bud-scar counts consistently underestimate the age of taller and older seedlings. But even when bud-scar estaimations of total seedling age had low accuracy, most recent yeras of growth are reliably identified. Because our juveniles range in age and height, I don't want to apply a correction factor across all of them. If we assume that our error rate was consistent across all of our sites (e.g., we underestimate 1 yr on all seedlings) and we're not going for TURLY ACCURATE establishment year but rather going for relative establishment and drivers of such -- then I don't need to apply a correction factor.
```{r setup, include=FALSE}

# mean_estab_yr_spp <- tripod %>%
#   group_by(., site) %>%
#   summarize_at(c("estab_yr"),
#                funs(mean(., na.rm=TRUE)))
# 
# median_estab_yr_spp <- tripod %>%
#   group_by(., site) %>%
#   summarize_at(c("estab_yr"),
#                funs(median(., na.rm=TRUE)))
# 
# min_estab_yr_spp <- tripod %>%
#   group_by(., site) %>%
#   summarize_at(c("estab_yr"),
#                funs(min(., na.rm=TRUE)))
# 
# max_estab_yr_spp <- tripod %>%
#   group_by(., site) %>%
#   summarize_at(c("estab_yr"),
#                funs(max(., na.rm=TRUE)))

ttl_by_spp <- tripod %>%
  filter(! spp == "SEEDLING") %>%
  group_by(., spp) %>%
  summarize_at(c("tpha"),
               funs(sum(., na.rm=TRUE)))

min_estab_yr_individ <- tripod %>%
  filter(! spp == "SEEDLING") %>%
  group_by(., site, spp) %>%
  summarize_at(c("estab_yr"),
               funs(min(., na.rm=TRUE)))

# For figure and comparison purposes, drop any NONEs and any Inf (i.e., don't have yr recorded)
drops <- c("NONE", "UNKN")
min_estab_yr_individ <-min_estab_yr_individ[! min_estab_yr_individ$spp %in% drops,]
drops <- Inf
min_estab_yr_individ <-min_estab_yr_individ[! min_estab_yr_individ$estab_yr %in% drops,]

data <- min_estab_yr_individ

# Boxplot of minimum establishment years by individual species (of all the individs sampled)

spp.names <- c("subalpine\nfir",
               "western\nlarch",
               "lodgepole\npine",
               "Engelmann\nspruce",
               "ponderosa\npine",
               "Douglas-\nfir")

estab_plot <- ggplot(data = data, aes(x = spp, y = estab_yr))
estab_plot + geom_boxplot() +
  labs(x = "Species", y = "Earliest year of establishment") + 
  scale_x_discrete(labels = spp.names) + 
  scale_y_continuous(breaks=2006:2016) +
  coord_flip() + 
  theme_caitlin()


# More interesting?
# estab_plot + geom_violin(fill="black") +
#   labs(x = "Species", y = "Earliest year of establishment") + 
#   scale_x_discrete(labels = spp.names) +
#   coord_flip() +
#   theme_caitlin()

currentDate <- Sys.Date()
ggsave(paste0("min_estab_yr_vs_spp_",currentDate,".png"), width = 5, height = 3)
dev.off()

# ANOVA to test difference between individual species minimum establishment years
mod.yr <- lm(estab_yr ~ spp, data = data)
anova(mod.yr)
aov <- aov(estab_yr ~ spp, data = data)

# OK, yes significant difference, but now need to do pair-wise comparison
TukeyHSD(aov)

# Or b/c not normal, use non-parametric Kruskal-Wallis
library(dunn.test)
kw <- kruskal.test(estab_yr ~ spp, data = data) ; kw
dunn <- dunn.test(x=data$estab_yr, g=data$spp, kw=TRUE)

remove(mod.yr, aov, kw, dunn)

# Maybe I should consider 2-way ANOVA to include min dist seed source...?
# Or it may be a sufficient conclusion to demonstrate that PICO is diff from all.

```


```{r FILL IN MISSING YRS}
(library(dplyr))
# Try capturing all the trees... fill in NAs at any given site with reps from that site
yrly_estab <- tripod %>%
  filter(! spp == "SEEDLING") %>%
  group_by(.,site, estab_yr, spp) %>%
  summarize_at(c("tpha"),
               funs(sum(., na.rm=TRUE)))

# Drop records (plots/quadrants) with no trees
yrly_estab$estab_yr <- as.factor(yrly_estab$estab_yr)
drops <- c("NONE")
yrly_estab <-yrly_estab[! yrly_estab$spp %in% drops,]


# drops are the records that have yr tbd; temp are records that have known yr
drops <- yrly_estab[is.na(yrly_estab$estab_yr),] %>%
  dplyr::rename(tbd = tpha)
temp <- yrly_estab[! is.na(yrly_estab$estab_yr),]

# in drops
# S-322-577 NA ABLA 25.46479 <- won't be able to distribute b/c no aged abla at that site
# S-332-720 NA ABLA 76.39437
# in temp
# S-231-304 2009 ABLA 25.46479
# S-321-502 2009 ABLA 25.46479
# S-332-720 2011 ABLA 50.9295 <- will be surrogate for unaged 720s above



# I want to fill in all the estab_yr tbds. Add new column that gives total of that species...
# with KNOWN years so I can then get proportion of aged trees of a given species, given yr.
temp1 <- temp %>%
  dplyr::group_by(., site, spp) %>%  
  dplyr::mutate(ttl = sum(tpha)) %>% # gives total of that spp at site across all yrs
  dplyr::mutate(prop = tpha/ttl) # gives prop of that yr for given spp of that spp at site
# I'll use that proportion of each-yr representation to assign non-aged trees (aka tbd, stored in drops)

# Gives table with all legit tpha records replicated for each drop at given site.
temp2 <- full_join(temp1, drops, by = "site")
# See that ALBA 2011 have corresponding ABLA (unaged) from drops
# See that ABLA 2009 have NO corresponding ABLA (unaged) from drops
# Make sure I don't lose either 2009 or 2011 that are actually measured!!
# Then somehow quantify the number of trees that remain unaged -- maybe add back in somehow?


# If spp from actually aged data (tpha) and non-aged data (tbd) match, then distribute tbd
# Letting tpha.new exist if I'm adding conspecific tpha and tbd. 
# OR letting tpha.new exist if there's no tbd -- just keep original tpha.
temp3 <- temp2 %>%
  mutate(add = ifelse(spp.x == spp.y, prop*tbd, 0)) %>% # if species match, compute # tbd to distribute
  mutate(tpha.new = ifelse((spp.x == spp.y | prop == 1), tpha + add, NA)) %>% 
  filter(!is.na(tpha.new))
# ABLA from 2009 and from 2011 remain (with new assigned unaged ones at site 720)

# Renames and selects relevant variabes; nixes duplicate spp&yr which had no tbd
temp4 <- temp3 %>%
  dplyr::rename(estab_yr = estab_yr.x, spp = spp.x) %>%
  dplyr::select(site, estab_yr, spp, tpha.new) %>% # dplyr select classes with MASS
  distinct()

# Take average across all sites of each species established in each year
data <- temp4 %>%
  group_by(estab_yr, spp) %>%
  summarize_at(c("tpha.new"),
               funs(mean(., na.rm=TRUE)))

# Now find all those drops that didn't have a home...
# ...when ALL indivds of that spp at that site were unaged
# anti_join gives all rows in x where no matching y, just keeping col from x
aged <- temp %>% dplyr::group_by(site, spp) %>% dplyr::summarize(n.aged=n())
ageless <- anti_join(drops, aged)
# These 18 observations have no age and no age to reference.
# Not too many so I'm NOT going to distribute those ages based on study area representation.
# Just need to add caveat below estab age figures.

remove(temp, temp1, temp2, temp3, temp4, drops, aged, ageless)

data.orig <- data

# I'll want to preserve all the years in each plot; n.b., 2016 isn't here b/c I've filtered germinants out of dataset
all_yrs <- as.data.frame(matrix(c("2006","2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015"), ncol=1, byrow=TRUE))
all_yrs <- all_yrs %>%
  dplyr::rename(estab_yr = V1)
yr_since <- (0:9)
```


```{r PLOTS OF ESTAB YR}
# Load climate data
source("Tripod_01g_ClimateData.v2.R")
# Potential climate variables to use as color bars.
# Not using annual b/c includes conditions from after growing season.
# Creating SPEI avg for select months means grouping, which drops attributes.
# Re-adding attributes drops names of new column.
# N.b., if I want to do temp or precip, change plot limits and for precip, do SUM not MEAN
# Also n.b., dplyr doesn't like un-tidy data so summarizing across rows just returned first value.
# Forcing it to use untidy data here by spelling out average:
var <- wwdt_spei %>%
  mutate(amjj = (m04 + m05 + m06 + m07)/4) %>%
  mutate(mjja = (m05 + m06 + m07 + m08)/4) %>%
  mutate(amjjas = (m04 + m05 + m06 + m07 + m08 + m09)/6) %>%
  mutate(year = rownames(wwdt_spei))
# write.csv(var, "summary_spei.csv")
# 
# var <- wwdt_temp %>%
#   mutate(amjj = (m04 + m05 + m06 + m07)/4) %>%
#   mutate(amjjas = (m04 + m05 + m06 + m07 + m08 + m09)/6) %>%
#   mutate(year = rownames(wwdt_temp))
# write.csv(var, "summary_temp.csv")
# 
# 
# var <- wwdt_precip %>%
#   mutate(amjj = (m04 + m05 + m06 + m07)) %>%
#   mutate(mjja = (m05 + m06 + m07 + m08)) %>%
#   mutate(amjjas = (m04 + m05 + m06 + m07 + m08 + m09)) %>%
#   mutate(year = rownames(wwdt_precip))
# write.csv(var, "summary_precip.csv")


var <- var %>%
  dplyr::filter(year == c(2006:2015)) %>%
  # dplyr::select(mjja)
  dplyr::select(amjj)
  
var <-c(var$amjj)

# Secondary axis (sec_axis()) must be 1:1 transformation of primary axis.
# Here, setting 100 to equal twice the max value of the tpha
# Figure out what to do with color guide, now turned off.

####ABLA#######################################################################
# ABLA 
# Select data
data <- data.orig[which(data.orig$spp == "ABLA"),]
data <- left_join(all_yrs, data, by = "estab_yr")
data$tpha.new[which(is.na(data$tpha.new))] <- 0

# I'll want cumulative line about twice as high as max annual estab.
# This creates cumulative proportions and *max*2 translates/stretches values to double max.
cumscaled <- (cumsum(data$tpha.new)/sum(data$tpha.new)) * (max(data$tpha.new) * 1.05)

# colors=c("red", "white", "blue") # too purply
colors=c("sandybrown", "white", "skyblue") 
# colors=c("#ef8a62", "#f7f7f7", "#67a9cf") # more red to blue

# Create plot with 3 elements: tpha each yr, cumulative curve, drought each yr
abla_plot <- ggplot(data, aes(x = estab_yr, y = tpha.new)) +
  geom_col(width = 0.95) + # puts tiny space btwn columns (vs. width = 1)
  geom_col(aes(y = -(0.05*max(data$tpha.new)), fill = var), width = 1) + # fill in space below y=axis with drought indicator; scaling by max of data so these will be same size across plots
  scale_fill_gradientn(colors=colors,
                       values = c(0,0.5,1),  # mid num gives where, relatively, to put white
                       guide = "colourbar",
                       name = "SPEI",
                       # limits = c(min(var), max(var)), # alt is to set these to c(min(var), max(var))
                       # breaks = c(min(var), 0, max(var)), # alt is to set these to c(min(var),0,  max(var))
                       # labels = c(min(var), 0, max(var))) +
                       limits = c(-1.5, 1.5), # alt is to set these to c(min(var), max(var))
                       breaks = c(-1.5, 0, 1.5), # alt is to set these to c(min(var),0,  max(var))
                       labels = c(-1.5, 0, 1.5)) +
  geom_line(aes(y = cumscaled, group =1), linetype = 1, size = 1) + # add cumulative line
  scale_x_discrete(expand = c(0, 0), drop=FALSE, labels = data$estab_yr) + # don't add L/R space to x-axis
  scale_y_continuous(expand = c(0,0), # don't add any space above/below y-axis
                     limits = c(-(0.05*max(data$tpha.new)), # add space below y for drought bar
                                max(data$tpha.new) * 1.15), # the 0.15 adds lil space above
                     # add secondary axis: must be 1:1 transformation of primary axis. Here, 2*max = 100%
                     sec.axis = sec_axis(~./(max(data$tpha.new) * 1.05/100), name = "Cumulative %")) +
  labs(x = NULL, y = "Density (stems/ha)") +
  theme_caitlin() + 
  theme(legend.position= "none") +
  # annotate with name of spp, set at top of graph (there's spaec b/c limit is  2.05)
  annotate("text", x =1, y = max(data$tpha.new) * 1.1, label = "subalpine fir", hjust= 0)
abla_plot
temp <- abla_plot
abla_plot <- ggplotGrob(abla_plot) # If I don't save as object, seems to get re-drawn and loses axes
currentDate <- Sys.Date()
ggsave(paste0("estab_abla_spei_amjj_",currentDate,".png"), width = 5, height = 3)
dev.off()



###LAOC########################################################################
# LAOC 
data <- data.orig[which(data.orig$spp == "LAOC"),]
data <- left_join(all_yrs, data, by = "estab_yr")
data$tpha.new[which(is.na(data$tpha.new))] <- 0
cumscaled <- (cumsum(data$tpha.new)/sum(data$tpha.new)) * (max(data$tpha.new) * 1.05)
laoc_plot <- ggplot(data, aes(x = estab_yr, y = tpha.new)) +
  geom_col(width = 0.95) + 
  geom_col(aes(y = -(0.05*max(data$tpha.new)), fill = var), width = 1) +
  scale_fill_gradientn(colors=c("sandybrown", "white", "skyblue"),
                       values = c(0,0.5,1), 
                       guide = "colourbar",
                       name = "SPEI",
                       limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5),
                       labels = c(-1.5, 0, 1.5)) +
  geom_line(aes(y = cumscaled, group =1), linetype = 1, size = 1) +
  scale_x_discrete(expand = c(0, 0), drop=FALSE, labels = data$estab_yr) +
  scale_y_continuous(expand = c(0,0), # 
                     limits = c(-(0.05*max(data$tpha.new)),
                                max(data$tpha.new) * 1.15),
                     sec.axis = sec_axis(~./(max(data$tpha.new) * 1.05/100), name = "Cumulative %")) +
  labs(x = NULL, y = "Density (stems/ha)") +
  theme_caitlin() +
  theme(legend.position= "none") +
  annotate("text", x = 1, y = max(data$tpha.new) * 1.1, label = "western larch", hjust= 0)
laoc_plot
laoc_plot <- ggplotGrob(laoc_plot)
currentDate <- Sys.Date()
ggsave(paste0("estab_laoc_spei_amjj_",currentDate,".png"), width = 5, height = 3)
dev.off()


###PICO########################################################################
# PICO 
data <- data.orig[which(data.orig$spp == "PICO"),]
data <- left_join(all_yrs, data, by = "estab_yr")
data$tpha.new[which(is.na(data$tpha.new))] <- 0
cumscaled <- (cumsum(data$tpha.new)/sum(data$tpha.new)) * (max(data$tpha.new) * 1.05)
pico_plot <- ggplot(data, aes(x = estab_yr, y = tpha.new)) +
  geom_col(width = 0.95) + 
  geom_col(aes(y = -(0.05*max(data$tpha.new)), fill = var), width = 1) +
  scale_fill_gradientn(colors=c("sandybrown", "white", "skyblue"),
                       values = c(0,0.5,1), 
                       guide = "colourbar",
                       name = "SPEI",
                       limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5),
                       labels = c(-1.5, 0, 1.5)) +
  geom_line(aes(y = cumscaled, group =1), linetype = 1, size = 1) +
  scale_x_discrete(expand = c(0, 0), drop=FALSE, labels = data$estab_yr) +
  scale_y_continuous(expand = c(0,0), # 
                     limits = c(-(0.05*max(data$tpha.new)),
                                max(data$tpha.new) * 1.15),
                     sec.axis = sec_axis(~./(max(data$tpha.new) * 1.05/100), name = "Cumulative %")) +
  labs(x = NULL, y = "Density (stems/ha)") +
  theme_caitlin() + 
  theme(legend.position= "none") +
  annotate("text", x = 1, y = max(data$tpha.new) * 1.1, label = "lodgepole pine", hjust= 0)
pico_plot
pico_plot <- ggplotGrob(pico_plot)
currentDate <- Sys.Date()
ggsave(paste0("estab_pico_spei_amjj_",currentDate,".png"), width = 5, height = 3)
dev.off()

###PIEN########################################################################
# PIEN 
data <- data.orig[which(data.orig$spp == "PIEN"),]
data <- left_join(all_yrs, data, by = "estab_yr")
data$tpha.new[which(is.na(data$tpha.new))] <- 0
cumscaled <- (cumsum(data$tpha.new)/sum(data$tpha.new)) * (max(data$tpha.new) * 1.05)
pien_plot <- ggplot(data, aes(x = estab_yr, y = tpha.new)) +
  geom_col(width = 0.95) + 
  geom_col(aes(y = -(0.05*max(data$tpha.new)), fill = var), width = 1) +
  scale_fill_gradientn(colors=c("sandybrown", "white", "skyblue"),
                       values = c(0,0.5,1), 
                       guide = "colourbar",
                       name = "SPEI",
                       limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5),
                       labels = c(-1.5, 0, 1.5)) +
  geom_line(aes(y = cumscaled, group =1), linetype = 1, size = 1) +
  scale_x_discrete(expand = c(0, 0), drop=FALSE, labels = data$estab_yr) +
  scale_y_continuous(expand = c(0,0), # 
                     limits = c(-(0.05*max(data$tpha.new)),
                                max(data$tpha.new) * 1.15),
                     sec.axis = sec_axis(~./(max(data$tpha.new) * 1.05/100), name = "Cumulative %")) +
  labs(x = NULL, y = "Density (stems/ha)") +
  theme_caitlin() + 
  theme(legend.position= "none") +
  annotate("text", x = 1, y = max(data$tpha.new) * 1.1, label = "Englemann spruce", hjust= 0)
pien_plot
pien_plot <- ggplotGrob(pien_plot)
currentDate <- Sys.Date()
ggsave(paste0("estab_pien_spei_amjj_",currentDate,".png"), width = 5, height = 3)
dev.off()


###PIPO########################################################################
# PIPO 
data <- data.orig[which(data.orig$spp == "PIPO"),]
data <- left_join(all_yrs, data, by = "estab_yr")
data$tpha.new[which(is.na(data$tpha.new))] <- 0
cumscaled <- (cumsum(data$tpha.new)/sum(data$tpha.new)) * (max(data$tpha.new) * 1.05)
pipo_plot <- ggplot(data, aes(x = estab_yr, y = tpha.new)) +
  geom_col(width = 0.95) + 
  geom_col(aes(y = -(0.05*max(data$tpha.new)), fill = var), width = 1) +
  scale_fill_gradientn(colors=c("sandybrown", "white", "skyblue"),
                       values = c(0,0.5,1), 
                       guide = "colourbar",
                       name = "SPEI",
                       limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5),
                       labels = c(-1.5, 0, 1.5)) +
  geom_line(aes(y = cumscaled, group =1), linetype = 1, size = 1) +
  scale_x_discrete(expand = c(0, 0), drop=FALSE, labels = data$estab_yr) +
  scale_y_continuous(expand = c(0,0), # 
                     limits = c(-(0.05*max(data$tpha.new)),
                                max(data$tpha.new) * 1.15),
                     sec.axis = sec_axis(~./(max(data$tpha.new) * 1.05/100), name = "Cumulative %")) +
  labs(x = NULL, y = "Density (stems/ha)") +
  theme_caitlin() + 
  theme(legend.position= "none") +
  annotate("text", x = 1, y = max(data$tpha.new) * 1.1, label = "ponderosa pine", hjust= 0)
pipo_plot
pipo_plot <- ggplotGrob(pipo_plot)
currentDate <- Sys.Date()
ggsave(paste0("estab_pipo_spei_amjj_",currentDate,".png"), width = 5, height = 3)
dev.off()

###PSME########################################################################
# PSME 
data <- data.orig[which(data.orig$spp == "PSME"),]
data <- left_join(all_yrs, data, by = "estab_yr")
data$tpha.new[which(is.na(data$tpha.new))] <- 0
cumscaled <- (cumsum(data$tpha.new)/sum(data$tpha.new)) * (max(data$tpha.new) * 1.05)
psme_plot <- ggplot(data, aes(x = estab_yr, y = tpha.new)) +
  geom_col(width = 0.95) + 
  geom_col(aes(y = -(0.05*max(data$tpha.new)), fill = var), width = 1) +
  scale_fill_gradientn(colors=c("sandybrown", "white", "skyblue"),
                       values = c(0,0.5,1), 
                       guide = "colourbar",
                       name = "SPEI",
                       limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5),
                       labels = c(-1.5, 0, 1.5)) +
  geom_line(aes(y = cumscaled, group =1), linetype = 1, size = 1) +
  scale_x_discrete(expand = c(0, 0), drop=FALSE, labels = data$estab_yr) +
  scale_y_continuous(expand = c(0,0), # 
                     limits = c(-(0.05*max(data$tpha.new)),
                                max(data$tpha.new) * 1.15),
                     sec.axis = sec_axis(~./(max(data$tpha.new) * 1.05/100), name = "Cumulative %")) +
  labs(x = NULL, y = "Density (stems/ha)") +
  theme_caitlin() + 
  theme(legend.position= "none") +
  annotate("text", x = 1, y = max(data$tpha.new) * 1.1, label = "Douglas-fir", hjust= 0)
psme_plot
temp <- psme_plot
psme_plot <- ggplotGrob(psme_plot)
currentDate <- Sys.Date()
ggsave(paste0("estab_psme_spei_amjj_",currentDate,".png"), width = 5, height = 3)
dev.off()

###LEGEND########################################################################
# Dummy plot to get legend
leg_plot <- ggplot(data, aes(x = estab_yr, y = tpha.new)) +
  geom_col(aes(y = -(0.05*max(data$tpha.new)), fill = var), width = 1) +
  scale_fill_gradientn(colors=c("sandybrown", "white", "skyblue"),
                       values = c(0,0.5,1), 
                       guide = "colourbar",
                       name = "SPEI",
                       limits = c(-1.5, 1.5),
                       breaks = c(-1.5, 0, 1.5),
                       labels = c(-1.5, 0, 1.5)) +
  theme(legend.direction = "horizontal")
        
leg_plot

#Extract Legend 
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

legend <- g_legend(leg_plot)
dev.off()
grid.draw(legend)

```


```{r PLOT ALL TOGETHER NOW}
# PLOT ALL TOGETHER NOW

# install.packages("gridExtra")
library(gridExtra)
# https://cran.r-project.org/web/packages/gridExtra/vignettes/arrangeGrob.html
?grid.arrange

# png('test.png',
#      width=1105,
#      height=791)
# grid.arrange(abla_plot, laoc_plot, pico_plot, pien_plot, pipo_plot, psme_plot,
#              ncol = 2)
# dev.off()
# ^ Recall all are saved as graphical objects (Grob) or else they get re-drawn each time,
# which results in loss of individual axes and cumulative lines.
# But how do I draw in the legend??

# Some suggest switching to more powerful gtable
library(gtable)
lh <- rbind(abla_plot, laoc_plot, pico_plot, size = "first")
rh <- rbind(pien_plot, pipo_plot, psme_plot, size = "first")
g <- cbind(lh, rh, size = "first")
grid.newpage()
grid.draw(g)

# grid.arrange(g, legend, ncol = 1) # legend space is way too big
library(cowplot)
currentDate <- Sys.Date()
png(paste0("estab_all_spei_amjj_",currentDate,".png"),
    width=900, height =643)
plot_grid(g, legend, ncol = 1, align = "v", rel_heights = c(3, 0.25))
dev.off()




# To use below, need to NOT change to Grobs. But then axes are messed again. >:(
# I'm sure I could pull legend before changing to Grob but moving on...
# grid_arrange_shared_legend <- function(...) {
#     plots <- list(...)
#     g <- ggplotGrob(plots[[1]] + theme(legend.position="bottom"))$grobs
#     legend <- g[[which(sapply(g, function(x) x$name) == "guide-box")]]
#     lheight <- sum(legend$height)
#     grid.arrange(
#         do.call(arrangeGrob, lapply(plots, function(x)
#             x + theme(legend.position="none"))),
#         legend,
#         ncol = 1,
#         heights = unit.c(unit(1, "npc") - lheight, lheight))
# }
# 
# grid_arrange_shared_legend(abla_plot, laoc_plot, pico_plot, pien_plot, pipo_plot, psme_plot)


# Alt, if I need to better algin axes
# if(!require(devtools)) install.packages("devtools")
# devtools::install_github("kassambara/ggpubr")
# library(ggpubr)
# ggarrange(abla_plot, laoc_plot, pico_plot, pien_plot, pipo_plot, psme_plot,
#           # labels = c("A", "B", "C", "D", "e", "F"),
#           ncol = 2, nrow = 3)


# Try multiplot (which is a function I had to define)
# multiplot(abla_plot, laoc_plot, pico_plot, pien_plot, pipo_plot, psme_plot, cols = 2)

# Alt lines, but they'll bow below zero and above max:
  # geom_smooth(aes(y=cumsum(tpha.new), group = 1), # 1 group only
  # se=FALSE, method= "loess", # no SE bands, use local regression smoother
  # span = 0.75) + # span controls the wiggliness: fraction of pts used to fit

```


```{r ESTAB YR VS CLIMATE PREP}
# N.b., to use the climate data below, can't have run ClimGrwthCORR b/c it drops some yrs.
# Plan: identify significant correlations for current and future months.
# But only use estab yr histograms with growing season drougtht bars below for display.
# N.b., I'm including all of the newly aged juveniles (per procedure above)

source("Tripod_01g_ClimateData.v2.R")

wwdt_temp <- wwdt_temp %>%
  mutate(next01 = lead(m01, 1),
         next02 = lead(m02, 1),
         next03 = lead(m03, 1),
         next04 = lead(m04, 1),
         next05 = lead(m05, 1),
         next06 = lead(m06, 1),
         next07 = lead(m07, 1),
         next08 = lead(m08, 1),
         next09 = lead(m09, 1),
         next10 = lead(m10, 1),
         next11 = lead(m11, 1),
         next12 = lead(m12, 1)) 
rownames(wwdt_temp) <- c(2006:2016) # lost rownames in last move

wwdt_precip <- wwdt_precip %>%
  mutate(next01 = lead(m01, 1),
         next02 = lead(m02, 1),
         next03 = lead(m03, 1),
         next04 = lead(m04, 1),
         next05 = lead(m05, 1),
         next06 = lead(m06, 1),
         next07 = lead(m07, 1),
         next08 = lead(m08, 1),
         next09 = lead(m09, 1),
         next10 = lead(m10, 1),
         next11 = lead(m11, 1),
         next12 = lead(m12, 1)) 
rownames(wwdt_precip) <- c(2006:2016) # lost rownames in last move


wwdt_spei <- wwdt_spei %>%
  mutate(next01 = lead(m01, 1),
         next02 = lead(m02, 1),
         next03 = lead(m03, 1),
         next04 = lead(m04, 1),
         next05 = lead(m05, 1),
         next06 = lead(m06, 1),
         next07 = lead(m07, 1),
         next08 = lead(m08, 1),
         next09 = lead(m09, 1),
         next10 = lead(m10, 1),
         next11 = lead(m11, 1),
         next12 = lead(m12, 1)) 
rownames(wwdt_spei) <- c(2006:2016) # lost rownames in last move

year <- c("2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")
temp <- wwdt_temp[year,]
precip <- wwdt_precip[year,]
spei <- wwdt_spei[year,]

#ABLA
abla <- data.orig %>%
  filter(spp == "ABLA")
abla$estab_yr <- as.numeric(as.character(abla$estab_yr))
abla <- as.data.frame(abla)
abla <- abla %>%
  add_row(estab_yr = 2006, spp = "ABLA", tpha.new = 0) %>%
  add_row(estab_yr = 2007, spp = "ABLA", tpha.new = 0) %>%
  add_row(estab_yr = 2008, spp = "ABLA", tpha.new = 0) %>%
  add_row(estab_yr = 2010, spp = "ABLA", tpha.new = 0) %>%
  add_row(estab_yr = 2012, spp = "ABLA", tpha.new = 0) %>%
  add_row(estab_yr = 2013, spp = "ABLA", tpha.new = 0) %>%
  add_row(estab_yr = 2014, spp = "ABLA", tpha.new = 0) %>%
  add_row(estab_yr = 2015, spp = "ABLA", tpha.new = 0) %>%
  arrange(estab_yr)

#LAOC
laoc <- data.orig %>%
  filter(spp == "LAOC")
laoc$estab_yr <- as.numeric(as.character(laoc$estab_yr))
laoc <- as.data.frame(laoc)
laoc <- laoc %>%
  add_row(estab_yr = 2006, spp = "LAOC", tpha.new = 0) %>%
  add_row(estab_yr = 2007, spp = "LAOC", tpha.new = 0) %>%
  add_row(estab_yr = 2008, spp = "LAOC", tpha.new = 0) %>%
  arrange(estab_yr)


#PICO
pico <- data.orig %>%
  filter(spp == "PICO")
pico$estab_yr <- as.numeric(as.character(pico$estab_yr))
pico <- as.data.frame(pico)


#PIEN
pien <- data.orig %>%
  filter(spp == "PIEN")
pien$estab_yr <- as.numeric(as.character(pien$estab_yr))
pien <- as.data.frame(pien)
pien <- pien %>%
  add_row(estab_yr = 2006, spp = "PIEN", tpha.new = 0) %>%
  add_row(estab_yr = 2007, spp = "PIEN", tpha.new = 0) %>%
  add_row(estab_yr = 2013, spp = "PIEN", tpha.new = 0) %>%
  arrange(estab_yr)


#PIPO
pipo <- data.orig %>%
  filter(spp == "PIPO")
pipo$estab_yr <- as.numeric(as.character(pipo$estab_yr))
pipo <- as.data.frame(pipo)


#PSME
psme <- data.orig %>%
  filter(spp == "PSME")
psme$estab_yr <- as.numeric(as.character(psme$estab_yr))
psme <- as.data.frame(psme)
psme <- psme %>% add_row(estab_yr = 2006,
                         spp = "PSME",
                         tpha.new = 0) %>%
  arrange(estab_yr)

```


```{r ESTAB YR VS CLIMATE CORR TEST}
# Pull correlation test values out for each spp, for each climate var, for each month

# create empty vectors to fill
spp <- NULL
var.names <- NULL
cors <- NULL
ps <- NULL
ts <- NULL
dfs <- NULL
months <- NULL

# Pick variable and set so that loop is easy to re-run.
# Can keep adding on to the above vectors then combine all
year <- c("2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015")
var1 <- wwdt_temp[year,] ; var1.name <- "temp"
var2 <- wwdt_precip[year,] ; var2.name <- "precip"
var3 <- wwdt_spei[year,] ; var3.name <- "spei"

# data <- abla
# data <- laoc
# data <- pico
# data <- pien
# data <- pipo
data <- psme

month.loop.ready <- c(1:24)
for(j in month.loop.ready){

  data$spp <- as.character(data$spp)
  
  correlation <- cor.test(data[,"tpha.new"], var1[,j]) # Pearson
  sp <- data[1,"spp"] ; spp <- c(spp, sp) # grab the species name
  var.names <- c(var.names, var1.name)
  month <- colnames(var1[j]) ; months <- c(months, month)
  cor <-correlation$estimate ; cors <- c(cors, cor) 
  p <-correlation$p.value ; ps <- c(ps, p)
  df<- correlation$parameter ; dfs<-c(dfs,df)
  t<-correlation$statistic ;  ts<-c(ts, t)

  correlation <- cor.test(data[,"tpha.new"], var2[,j]) 
  sp <- paste0(data[1,2]) ; spp <- c(spp, sp) # grab the species name
  var.names <- c(var.names, var2.name)
  month <- colnames(var2[j]) ; months <- c(months, month)
  cor <-correlation$estimate ; cors <- c(cors, cor) 
  p <-correlation$p.value ; ps <- c(ps, p)
  df<- correlation$parameter ; dfs<-c(dfs,df)
  t<-correlation$statistic ;  ts<-c(ts, t)

  correlation <- cor.test(data[,"tpha.new"], var3[,j]) 
  sp <- paste0(data[1,2]) ; spp <- c(spp, sp) # grab the species name
  var.names <- c(var.names, var3.name)
  month <- colnames(var3[j]) ; months <- c(months, month)
  cor <-correlation$estimate ; cors <- c(cors, cor) 
  p <-correlation$p.value ; ps <- c(ps, p)
  df<- correlation$parameter ; dfs<-c(dfs,df)
  t<-correlation$statistic ;  ts<-c(ts, t)
}



###################################################################
### ONLY MAKE NEW DATAFRAME AFTER RUNNING ALL DESIRED VARIABLES ###
sp <- spp
var <- var.names
month <- months
cor <- cors
p.value <- ps
dfs <- dfs
t.stat <- ts
all_corrs <- data.frame(sp, var, month, cor, p.value, dfs, t.stat)

write.csv(all_corrs, "correlations.spp.estabyr.csv")
# ^ should really consider dropping ABLA, LAOC, PIEN (lotsa zeros)

library(corrplot)
cor <- cor(abla[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(laoc[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(pico[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(pien[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(pipo[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(psme[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
```




```{r TESTING GROUNDS -- DOES NEXT YEAR MATTER???}

source("Tripod_01g_ClimateData.v2.R")

wwdt_temp <- wwdt_temp %>%
  mutate(next01 = lead(m01, 1),
         next02 = lead(m02, 1),
         next03 = lead(m03, 1),
         next04 = lead(m04, 1),
         next05 = lead(m05, 1),
         next06 = lead(m06, 1),
         next07 = lead(m07, 1),
         next08 = lead(m08, 1),
         next09 = lead(m09, 1),
         next10 = lead(m10, 1),
         next11 = lead(m11, 1),
         next12 = lead(m12, 1)) 
rownames(wwdt_temp) <- c(2006:2016) # lost rownames in last move

wwdt_precip <- wwdt_precip %>%
  mutate(next01 = lead(m01, 1),
         next02 = lead(m02, 1),
         next03 = lead(m03, 1),
         next04 = lead(m04, 1),
         next05 = lead(m05, 1),
         next06 = lead(m06, 1),
         next07 = lead(m07, 1),
         next08 = lead(m08, 1),
         next09 = lead(m09, 1),
         next10 = lead(m10, 1),
         next11 = lead(m11, 1),
         next12 = lead(m12, 1)) 
rownames(wwdt_precip) <- c(2006:2016) # lost rownames in last move


wwdt_spei <- wwdt_spei %>%
  mutate(next01 = lead(m01, 1),
         next02 = lead(m02, 1),
         next03 = lead(m03, 1),
         next04 = lead(m04, 1),
         next05 = lead(m05, 1),
         next06 = lead(m06, 1),
         next07 = lead(m07, 1),
         next08 = lead(m08, 1),
         next09 = lead(m09, 1),
         next10 = lead(m10, 1),
         next11 = lead(m11, 1),
         next12 = lead(m12, 1)) 
rownames(wwdt_spei) <- c(2006:2016) # lost rownames in last move



cor <- cor(abla[,"tpha.new"], wwdt_temp[year,]) %>% corrplot
cor <- cor(laoc[,"tpha.new"], wwdt_temp[year,]) %>% corrplot
cor <- cor(pico[,"tpha.new"], wwdt_temp[year,]) %>% corrplot
cor <- cor(pien[,"tpha.new"], wwdt_temp[year,]) %>% corrplot
cor <- cor(pipo[,"tpha.new"], wwdt_temp[year,]) %>% corrplot
cor <- cor(psme[,"tpha.new"], wwdt_temp[year,]) %>% corrplot

cor <- cor(abla[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(laoc[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(pico[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(pien[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(pipo[,"tpha.new"], wwdt_precip[year,]) %>% corrplot
cor <- cor(psme[,"tpha.new"], wwdt_precip[year,]) %>% corrplot

cor <- cor(abla[,"tpha.new"], wwdt_spei[year,]) %>% corrplot
cor <- cor(laoc[,"tpha.new"], wwdt_spei[year,]) %>% corrplot
cor <- cor(pico[,"tpha.new"], wwdt_spei[year,]) %>% corrplot
cor <- cor(pien[,"tpha.new"], wwdt_spei[year,]) %>% corrplot
cor <- cor(pipo[,"tpha.new"], wwdt_spei[year,]) %>% corrplot
cor <- cor(psme[,"tpha.new"], wwdt_spei[year,]) %>% corrplot

```